Supplementary Materials for \"COATi: statistical pairwise alignment of protein coding sequences\

### 2024-02-05 alignpair_letter.tex
Complete RAC revisions. Version sent to Ziqi and Juan.

\begin{abstract}
\noindent
Sequence alignment is an essential method in bioinformatics and the basis of many analyses, including phylogenetic inference, ancestral sequence reconstruction, and gene annotation. Sequence artifacts and errors made in alignment reconstruction can impact downstream analyses leading to erroneous conclusions in comparative and functional genomic studies. For example, abiological frameshifts and early stop codons are common artifacts found in protein coding sequences that have been annotated in reference genomes. While such errors are eventually fixed in the reference genomes of model organisms, many genomes used by researchers contain these artifacts, and researchers often discard large amounts of data in comparative genomic studies to prevent artifacts from impacting results. To address this need, we present COATi, a statistical, codon-aware pairwise aligner that supports complex insertion-deletion models and can handle artifacts present in genomic data. COATi allows users to reduce the amount of discarded data while generating more accurate sequence alignments.
\end{abstract}

@@

% Figures are 86 mm or 178 mm wide
\begin{figure}[h!]
    \centering%
    \includegraphics[scale=1]{figures/fig-aln.pdf}
    \par
    \caption{
        Standard algorithms produce suboptimal alignments.
        (a) shows the true alignment of an ancestor sequence (A) and a descendant sequence (D).
        (b--d) are the results of different aligners. Nucleotide mismatches are highlighted in red.---Notably, COATi is the only aligner able to retrieve the biological alignment in this example.---%
        %
        Indels in protein-coding sequences can be classified as having one of three difference phases and being one of two different types.
        Phases refer the the location of the gap with respect to the reading frame, while types refer to the consequence of the indel.
        Phase-1, phase-2, and phase-3 indels are shown in purple, orange, and gray respectively.
        Additionally, the orange indel is type-II (an amino-acid indel plus an amino-acid change) while the purple indel is type-I (an amino-acid indel only). The difference between in-frame and frameshift indels is not displayed.
        }
    \label{fig:aln}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{figures/fig-fst-base-calling-new.pdf}
    \caption[FSTs more the generation of sequences]{%
    Finite state transducers (FSTs) model the generation of an output sequences based on an input sequence.
    %
    (a) A graph of a probabilistic FST \citep{cotterell-etal-2014-stochastic} for base-calling errors using a Mealy-machine architecture, where parameter $u$ is the error rate. This graph contains two states (S and M) connected by arcs, with labels ``input symbols~:~output symbols~/~weight''. Arcs consume symbols from the input sequence and emit symbols to the output sequence. Weights describe the probability that an arc is taken given the input symbols. Epsilon (ε) is a special symbol denoting that no symbols were either consumed or emitted.
    %
    (b) An FST for matching sequences against ambiguous nucleotides (N). This FST is not a true probabilistic FST and cannot be used to simulate output sequences since it is missing a parameter to control how often Ns are added to the output sequence.
    %
    (c) An FST that results from the composition ($∘$ operation) of the Error FST with the Ambiguity FST. As with (b), this composed FST cannot be used to simulate output sequences; however, it does properly weight the ambiguous nucleotide N as representing any other symbol.
    %
    %Let $x$ be a symbol in the input of (a), $y$ be a symbol in the output of (b), and $z$ be a symbol in both the output of (a) and the input of (b). An arc in the composed FST (c) has an input symbol $x$, an output symbol $y$, and a weight that is the combination over $z$ of the aggregate weights of all $x$-$z$-$y$ arc pairs.
    %
    }\label{fig:base-calling}
\end{figure}

@@

Since codon substitution is the first process in COATi's model, the input sequence to the COATi FST must be compatible with a codon-substitution model, i.e.\ be a multiple of three nucleotides and not contain any ambiguous symbols or stop codons. The phases, reading frames, and amino-acid contexts of alignment columns is determined by the input sequence, and better alignments will be generated if the input is of high quality and free of artifacts. Depending on context, we may refer to the input sequence as the ``ancestral'' or ``reference'' sequence. In contrast, the output sequence must be compatible with the ambiguity FST, can be of any length, and can contain any nucleotides or ``N''. This allows COATi to align lower-quality sequences that may contain artifacts against a high-quality reference sequence. We may refer to the output sequence as the ``descendant'', ``non-reference'', or ``query'' sequence. The choice of which sequence is the input sequence and which sequence is the output sequence is left up to the user.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{figures/fig-fst-coati.pdf}
\caption{The COATi FST is built from simpler FSTs via composition.
(a) The substitution FST encodes a $61 \times 61 $ codon substitution model with 3721 arcs from S to M. These arcs consume three nucleotides from the input tape and emit three nucleotides to the output tape. The weight of each arc is a conditional probability derived from a codon substitution model. See Fig.~\ref{fig:base-calling} for more details about reading this graph.
%
(b) The indel FST allows for insertions (H to I) and deletions (C to D). Here $g$ is the ``gap opening parameter'' and $e$ is the ``gap extension'' parameter.
Insertion arcs are weighted according to the codon model's stationary distribution of nucleotides, and deletion arcs have a weight of 1. This FST is structured such that if insertions and deletions are contiguous, insertions will precede deletions.
%
(c) The COATi FST is derived via composition from the codon substitution, indel, error, and ambiguity FSTs.
}
\label{fig:coati-fst}
\end{figure}

In order to use the COATi FST to align an output sequence against an input sequence, we first convert each sequence into an acceptor, represented as a linear transducer where the input and output symbols of each transition are identical and each transition represents one nucleotide of a sequence \citep{allauzen2007openfst}. By composing the input and output acceptors with the COATi FST, we generate a transducer of all possible alignments of the two sequences. Any path through this FST represents a pairwise alignment, while the shortest path (by weight) corresponds to the best alignment. If more that one optimal alignment exists, ties are broken according to the implementation of the shortest-path algorithm. All FST operations in COATi, including model development, composition, search for the shortest path, and other optimization algorithms, are performed using the C++ openFST library \citep{allauzen2007openfst}. An example of an FST-based alignment can be found in Supplementary Materials Figure 1.

@@

Intuitively, $d_{seq}$ ranges between zero and one and can be interpreted as the fraction of nucleotides in the sequence pair that are aligned differently between estimated and benchmark alignments. This metric summarizes each alignment by building homology sets for each nucleotide in the sequence pair. Briefly, if an alignment column contains position $i$ from the first sequence and position $j$ from the second sequence, then $\{j\}$ is the homology set for position $i$ in the first sequence and $\{i\}$ is the homology set for position $j$ in the second sequence. If a position is aligned against a gap, then its homology set is simply $\{\text{gap}\}$. This treats all gaps equally as the location of the gap is not recorded. Finally, $d_{seq}$ is calculated  between an estimated and a benchmark alignment as the average, normalized Hamming distance between the homology sets for each nucleotide in the sequence pair. See \cite{metrics_blackburne_whelan_2011} for further details.

@@

Using $d_{seq}$ we quantified the number of perfect, imperfect and best alignments each aligner produced. We define perfect alignments as alignments with a distance of zero to the benchmark alignment ($d_{seq} = 0$) or any alignment that is evolutionary equivalent to an alignment with a distance of zero. Equivalency was included in our definition of perfect alignments to prevent the manner by which aligners break ties between multiple optimal alignments from influencing our results. Evolutionary equivalence was determined by scoring alignments using COATi's marginal model. Any alignment that had a score which matched the score of its benchmark alignment was considered perfect even if its distance to the benchmark was greater than zero.
%
Imperfect alignments are defined as alignments that are not perfect when another method successfully produces a perfect alignment for the same pair of sequences.
%
Best alignments are the alignments that the lowest distance $d_{seq}$ to the true alignment, including ties.
%
Taken together, these three statistics not only allow a direct comparison of aligners but also expose instances where all aligners fall short of achieving a perfect result.

@@

As expected, COATi was not able to align 23 sequence pairs and COATi-rev was not able to align 193 sequence pairs. PRANK was not able to align 24 sequence pairs. MAFFT, MACSE, and ClustalΩ were able to align all sequence pairs, although the latter received some help from its wrapper script. Additionally, MAFFT produced alignments with columns that contained only gaps, which were removed by its wrapper script.
A total of 10296 sequence pairs were aligned equivalently by all five methods, including 8261 sequence pairs that did not contain any gaps, 664 that had identical alignments, and another 1371 that had identical scores.

@@

We have developed a statistical alignment program using finite-state transducers called COATi. On empirical data, COATi produced alignments that provided reasonable biological inferences. On semi-empirical, simulated benchmarks COATi was more accurate than the other aligners/alignment strategies that we tested: ClustalΩ's amino-acid model, MACSE's amino-acid away nucleotide model, MAFFT's DNA model, and PRANK's codon model. COATi's advantage over other aligners in our tests is at least partially due to its default parameters fitting the human-gorilla coding sequences better than the other aligner's default settings. Different data sets may show different results. However, because COATi uses a statistical model with biologically meaningful parameters, users can easily optimize COATi by using parameters estimated by previous studies or estimating parameters themselves \citep[e.g.][]{zhu2022profiling}.

COATi's primary disadvantage is that its FST model was slower and consumed more resources than other methods. However, COATi's marginal model produces results similar to the FST model, was the fastest method tested, and consumed reasonable amounts of memory \citep{garcia2023dissertation}. COATi is also not a symmetric pairwise aligner, as reference sequences are more constrained than non-reference sequences.
%
In order to test whether the choice of a reference sequence impact's COATi's results, we aligned our empirical sequence pairs with COATi using gorilla as the reference. We found no biologically or statistically significant differences between the results. The most noticeable difference was that with gorilla as the reference COATi rejected more sequence pairs due to the presence of ambiguous nucleotides in gorilla sequences.

After COATi, PRANK's codon model was the second most accurate alignment method. Like COATi, PRANK also uses statistical models and supports aligning protein-coding sequences using codon models developed for phylogenetics. However, PRANK's codon model only permits phase-3 gaps and is very sensitive to frameshift artifacts. In fact, PRANK refuses to align any sequence using a codon model if its length is not a multiple of three. Similarly, ClustalΩ amino-acid model also only permits phase-3 gaps and is very sensitive to frameshift artifacts; however, it was the worst performing protocol that we tested. It had the highest average alignment error and also had difficulties correctly identifying positive selection and estimating evolutionary distances.  --- It even performed poorly on its own gap patterns. --- ClustalΩ's relatively poor performance likely derives from its default assumptions being a poor fit for our data set and from the loss of information that occurred when we translated protein-coding DNA sequences into protein sequences.

@@

We can look at distances between alignments generated by different aligners to understand the diversity of approaches use by our alignment protocols. Aside from the cluster of different COATi methods, we detected no clusters in our analysis. Different aligners tended to spread out from one another and in different directions from the benchmark. This indicates that each protocol approached alignment inference from a different section of ``aligner space''. While each aligner, except ClustalΩ, performed well on its own gap patterns, MACSE also did well on ClustalΩ and MAFFT gap patterns, and PRANK also did well on ClustalΩ patterns. COATi did well on all patterns.

>>>>

\begin{abstract}
\noindent
Sequence alignment is an essential method in bioinformatics and the basis of many analyses, including phylogenetic inference, ancestral sequence reconstruction, and gene annotation. Sequence artifacts and errors made during alignment reconstruction can impact downstream analyses leading to erroneous conclusions in comparative and functional genomic studies. For example, abiological frameshifts and early stop codons are common artifacts found in protein coding sequences that have been annotated in reference genomes. While such errors are eventually fixed in the reference genomes of model organisms, many genomes used by researchers contain these artifacts, and researchers often discard large amounts of data in comparative genomic studies to prevent artifacts from impacting results. To address this need, we present COATi, a statistical, codon-aware pairwise aligner that supports complex insertion-deletion models and can handle artifacts present in genomic data. COATi allows users to reduce the amount of discarded data while generating more accurate sequence alignments.
\end{abstract}

@@

% Figures are 86 mm or 178 mm wide
\begin{figure}[h!]
    \centering%
    \includegraphics[scale=1]{figures/fig-aln.pdf}
    \par
    \caption{
        Standard algorithms produce suboptimal alignments.
        (a) shows the true alignment of an ancestor sequence (A) and a descendant sequence (D).
        (b--d) are the results of different aligners. Nucleotide mismatches are highlighted in red.---Notably, COATi is the only aligner able to retrieve the biological alignment in this example.---%
        %
        Indels in protein-coding sequences can be classified as having one of three difference phases and being one of two different types.
        Phases refer the the location of the gap with respect to the reading frame, while types refer to the consequence of the indel.
        Phase-1, phase-2, and phase-3 indels are shown in blue, orange, and green respectively.
        Additionally, the orange indel is type-II (an amino-acid indel plus an amino-acid change) while the blue indel is type-I (an amino-acid indel only). The difference between an in-frame and a frameshift indel is not displayed.
        }
    \label{fig:aln}
\end{figure}

@@

\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{figures/fig-fst-base-calling-new.pdf}
    \caption[FSTs more the generation of sequences]{%
    Finite state transducers (FSTs) model the generation of an output sequence based on an input sequence.
    %
    (a) A graph of a probabilistic FST \citep{cotterell-etal-2014-stochastic} for base-calling errors using a Mealy-machine architecture, where parameter $u$ is the error rate. This graph contains two states (S and M) connected by arcs, with labels ``input symbols~:~output symbols~/~weight''. Arcs consume symbols from the input sequence and emit symbols to the output sequence. Weights describe the probability that an arc is taken given the input symbols. Epsilon (ε) is a special symbol denoting that no symbols were either consumed or emitted.
    %
    (b) An FST for matching sequences against ambiguous nucleotides (N). 
    (c) An FST that results from the composition ($∘$ operation) of the Error FST with the Ambiguity FST.
    %
    %Let $x$ be a symbol in the input of (a), $y$ be a symbol in the output of (b), and $z$ be a symbol in both the output of (a) and the input of (b). An arc in the composed FST (c) has an input symbol $x$, an output symbol $y$, and a weight that is the combination over $z$ of the aggregate weights of all $x$-$z$-$y$ arc pairs.
    %
    }\label{fig:base-calling}
\end{figure}

@@

COATi FST is not a true probabilistic FST and cannot be used as-is to simulate output sequences based on an input sequence. This is because it is missing a parameter to control how often Ns are added to the output sequence. This is design feature to allow COATi FST to properly weight ambiguous nucleotides as representing any other symbol. In addition, COATi's indel FST (Fig.~\ref{fig:coati-fst}b) implicitly assumes that match states occur both immediately before and immediately after the alignment. This allows COATi FST to assign the same weight to gaps at the beginning and end of an alignment but also introduces a normalization constant (not shown) to reflect that mass is lost by not allowing the indel FST to terminate from node C. While this normalization constant is not needed to find the most likely alignment or to sample alignments, it can be calculated using a Markov model that considers only the transitions between states M and D in the indel FST (Fig.~\ref{fig:coati-fst}b) and ignoring any states in between:
\[
\frac{1-e + g(e-g)^{n+1}}{1-e+g} \approx \frac{1-e}{1-e+g} 
\]
where $g$ is the gap opening parameter, $e$ is the gap-extension parameter, and $n$ is the length of the input sequence. The exponent has a $+1$ due to the presence of an ``immortal match'' immediately after the input sequence.

Since codon substitution is the first process in COATi's model, the input sequence to the COATi FST must be compatible with a codon-substitution model and be a multiple of three nucleotides and not contain any ambiguous symbols or stop codons. The phases, reading frames, and amino-acid contexts of alignment columns is determined by the input sequence, and better alignments will be generated if the input is of high quality and free of artifacts. Depending on context, we may refer to the input sequence as the ``ancestral'' or ``reference'' sequence. In contrast, the output sequence must be compatible with the ambiguity FST, can be of any length, and can contain any nucleotides or ``N''. This allows COATi to align lower-quality sequences that may contain artifacts against a high-quality reference sequence. We may refer to the output sequence as the ``descendant'', ``non-reference'', or ``query'' sequence. The choice of which sequence is the input sequence and which sequence is the output sequence is left up to the user.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{figures/fig-fst-coati.pdf}
\caption{The COATi FST is built from simpler FSTs via composition.
(a) The substitution FST encodes a $61 \times 61 $ codon substitution model with 3721 arcs from S to M. These arcs consume three nucleotides from the input tape and emit three nucleotides to the output tape. The weight of each arc is a conditional probability derived from a codon substitution model. See Fig.~\ref{fig:base-calling} for more details about reading this graph.
%
(b) The indel FST allows for insertions (H to I) and deletions (C to D). Here $g$ is the gap-opening parameter and $e$ is the gap-extension parameter.
Insertion arcs are weighted according to the codon model's stationary distribution of nucleotides, and deletion arcs have a weight of 1. This FST is structured such that if insertions and deletions are contiguous, insertions will precede deletions.
%
(c) The COATi FST is derived via composition from the codon substitution, indel, error, and ambiguity FSTs.
}
\label{fig:coati-fst}
\end{figure}

In order to use the COATi FST to align an output sequence against an input sequence, we first convert each sequence into an acceptor, represented as a linear transducer where the input and output symbols of each transition are identical and each transition represents one nucleotide of a sequence \citep{allauzen2007openfst}. By composing the input and output acceptors with the COATi FST, we generate a transducer of all possible alignments of the two sequences. Any path through this FST represents a pairwise alignment, while the shortest path (by weight) corresponds to the best alignment. If more that one optimal alignment exists, ties are broken according to the implementation of the shortest-path algorithm. All FST operations in COATi, including model development, composition, search for the shortest path, and other optimization algorithms, are performed using the C++ openFST library \citep{allauzen2007openfst}. An example of an FST-based alignment can be found in Supplementary Materials Figure S1.

@@

Intuitively, $d_{seq}$ ranges between zero and one and can be interpreted as the fraction of nucleotides in the sequence pair that are aligned differently between estimated and benchmark alignments. This metric summarizes each alignment by building homology sets for each nucleotide in the sequence pair. Briefly, if an alignment column contains position $i$ from the first sequence and position $j$ from the second sequence, then $\{j\}$ is the homology set for position $i$ in the first sequence and $\{i\}$ is the homology set for position $j$ in the second sequence. If a position is aligned against a gap, then its homology set is simply $\{\text{gap}\}$. This treats all gaps equally as the location of the gap is not recorded. Finally, $d_{seq}$ is calculated  between an estimated and a benchmark alignment as the average, normalized Hamming distance between the homology sets for each nucleotide in the sequence pair \citep{metrics_blackburne_whelan_2011}.

@@

Using $d_{seq}$ we quantified the number of perfect, imperfect, and best alignments each aligner produced. We define perfect alignments as alignments with a distance of zero to the benchmark alignment ($d_{seq} = 0$) or any alignment that is evolutionary equivalent to an alignment with a distance of zero. We included equivalency in our definition of perfect alignments to prevent the manner by which aligners break ties from influencing our results. Evolutionary equivalence was determined by scoring alignments using COATi's marginal model. Any alignment that had a score which matched the score of its benchmark alignment was considered perfect even if its distance to the benchmark was greater than zero.
%
Imperfect alignments are defined as alignments that are not perfect when another method successfully produces a perfect alignment for the same pair of sequences.
%
Best alignments are the alignments that the lowest distance $d_{seq}$ to the true alignment, including ties and equivalent alignments.
%
Taken together, these three statistics not only allow a direct comparison of aligners but also expose instances where all aligners fall short of achieving a perfect result.

@@

As expected, COATi was not able to align 23 sequence pairs and COATi-rev was not able to align 193 sequence pairs. PRANK was not able to align 24 sequence pairs. MAFFT, MACSE, and ClustalΩ were able to align all sequence pairs, although the latter received some help from its wrapper script. Additionally, MAFFT produced alignments with columns that contained only gaps, which were removed by its wrapper script.
A total of 10,296 sequence pairs were aligned equivalently by all five methods, including 8,261 sequence pairs that did not contain any gaps, 664 that had identical alignments, and another 1,371 that had identical scores.

@@

We have developed a statistical alignment program using finite-state transducers called COATi. On empirical data, COATi produced alignments that provided reasonable biological inferences. On semi-empirical, simulated benchmarks COATi was more accurate than the other aligners/alignment strategies that we tested: ClustalΩ's amino-acid model, MACSE's amino-acid aware nucleotide model, MAFFT's DNA model, and PRANK's codon model. Unlike standard codon or amino-acid alignment strategies, COATi supports aligning protein-coding sequences using all three phases of gaps. Consistent with \cite{taylor2004occurrence} and \cite{zhu2022profiling}, our COATi results estimate that only 41\% indels in protein-coding sequences between humans and gorillas are phase-3 indels. This indicates that alignment strategies that only support phase-3 gaps in protein-coding sequences are sub-optimally aligning the 59\% of indels that are phase-1 or phase-2.

In additional to supporting all three phases of gaps, COATi's other advantage over competing aligners in our tests is at least partially due to its default parameters fitting the human-gorilla coding sequences better than the other aligner's default settings. Different data sets may show different results. However, because COATi uses a statistical model with biologically meaningful parameters, users can easily optimize COATi by using parameters estimated by previous studies or estimating parameters themselves \citep[e.g.][]{zhu2022profiling}.

COATi's primary disadvantage is that its FST model was slower and consumed more resources than other methods. However, COATi's marginal model produces results similar to the FST model, was the fastest method tested, and consumed reasonable amounts of memory (Supplementary Materials and \citealt{garcia2023dissertation}).
%
COATi is not a symmetric pairwise aligner, as reference sequences are more constrained than non-reference sequences. In order to test whether the choice of a reference sequence impact's COATi's results, we aligned our empirical sequence pairs with COATi using gorilla as the reference. We found no biologically or statistically significant differences between the results. The most noticeable difference was that, with gorilla as the reference, COATi rejected more sequence pairs due to the presence of ambiguous nucleotides in gorilla sequences.

After COATi, PRANK's codon model was the second most accurate alignment method. Like COATi, PRANK also uses statistical models and supports aligning protein-coding sequences using codon models developed for phylogenetics. Both COATi and PRANK had reasonable distributions of evolutionary distances that lacked large right tails. This is likely because COATi and PRANK produced better alignments when tasked with aligning sequence pairs that contained non-orthologous exons. However, in contrast with COATi, PRANK's codon model only permits phase-3 gaps and is very sensitive to frameshift artifacts. In fact, PRANK refuses to align any sequence using a codon model if its length is not a multiple of three.

ClustalΩ's amino-acid model also only permits phase-3 gaps and is very sensitive to frameshift artifacts; however, it was the worst performing protocol that we tested. It had the highest average alignment error and also had difficulties correctly identifying positive selection and estimating evolutionary distances.  --- It even performed poorly on its own gap patterns. --- ClustalΩ's relatively poor performance likely derives from its default assumptions being a poor fit for our data set and also from the loss of information that occurred when we translated protein-coding DNA sequences into protein sequences.

@@

We can look at distances between alignments generated by different aligners to understand the diversity of approaches use by our alignment protocols. Aside from the cluster of different COATi methods, we detected no other clusters in our analysis (Supplemental Materials Fig.~S6). Different aligners tended to spread out from one another and in different directions from the benchmark. This indicates that each protocol that we utilized approached alignment inference from a different section of ``aligner space''. While each aligner, except ClustalΩ, performed well on its own gap patterns, MACSE also did well on ClustalΩ and MAFFT gap patterns, and PRANK also did well on ClustalΩ patterns. COATi did well on all patterns.

### 2024-02-05 figures/fig-fst-base-calling-new.tex
Complete RAC revisions. Version sent to Ziqi and Juan.

@@

\tikzset{
  end/.style={fill=none, text=black, inner sep=1pt},
  block/.style={circle,on grid,fill=colorBu,text=white,minimum width=2em,node distance=4em},
  line/.style={draw, -Stealth, thick},
  lab/.style={text = black, inner sep=1pt},
  title/.style={text = black, font=\bfseries}
}

>>>>

\definecolor{TolHighContrastBlue}{RGB}{0,68,136}
\definecolor{TolVibrantBlue}{RGB}{0,119,187}

@@

\tikzset{
  end/.style={fill=none, text=black, inner sep=1pt},
  block/.style={circle,on grid,fill=TolVibrantBlue,text=white,minimum width=2em,node distance=4em},
  line/.style={draw, -Stealth, thick},
  lab/.style={text = black, inner sep=1pt},
  title/.style={text = black, font=\bfseries}
}

### 2024-02-05 figures/fig-fst-coati.tex
Complete RAC revisions. Version sent to Ziqi and Juan.

@@

\draw[white,line width=2pt] ($(w)!0.245!(i)$) -- +(0,4pt);
\draw[white,line width=2pt] ($(w)!0.385!(i)$) -- +(0,4pt);
\draw[white,line width=2pt] ($(w)!0.525!(i)$) -- +(0,4pt);
\draw[white,line width=2pt] ($(w)!0.665!(i)$) -- +(0,4pt);

@@

\draw[line] (s) to[out=225,in=0] (suba) to[out=180,in=350] (m);
\draw[line] (s) to[out=240,in=0] (subc) to[out=180,in=335] (m);
\draw[line] (s) to[out=255,in=0] (subg) to[out=180,in=320] (m);
\draw[line] (s) to[out=270,in=0] (subt) to[out=180,in=305] (m);

>>>>

\definecolor{TolVibrantBlue}{RGB}{0,119,187}

@@

@@

\draw[white,line width=2pt] (suba -| w) +(0,-2pt) -- +(0,2pt);
\draw[white,line width=2pt] (subc -| w) +(0,-2pt) -- +(0,2pt);
\draw[white,line width=2pt] (subg -| w) +(0,-2pt) -- +(0,2pt);
\draw[white,line width=2pt] (subt -| w) +(0,-2pt) -- +(0,2pt);

\draw[line] (s) to[out=225,in=0] (suba) to ($(suba)+(-20mm,0mm)$) to[out=180,in=330] (m);
\draw[line] (s) to[out=240,in=0] (subc) to ($(subc)+(-18mm,0mm)$) to[out=180,in=315] (m);
\draw[line] (s) to[out=255,in=0] (subg) to ($(subg)+(-15mm,0mm)$) to[out=180,in=300] (m);
\draw[line] (s) to[out=270,in=0] (subt) to ($(subt)+(-10mm,0mm)$) to[out=180,in=285] (m);
